---
title: "Introduction to GiottoVisuals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to GiottoVisuals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Description of the module GiottoVisuals

GiottoVisuals contains the main plotting functions of Giotto Suite. 

For full examples of the usage of Giotto visualizations visit http://giottosuite.com

## Script Organization by Prefixes:

aux_ - auxiliary and meta functionality
- aux_output.R - plot output handling
- aux_save.R - plot saving
- aux_defaults.R - set plotting defaults
- aux_visuals.R - general auxiliary plotting functions

plot_ - general plotting functions organized by type of plot
- plot_scatter.R - scatter plots
- plot_violin.R - violin plots
- plot_heatmap.R - heatmaps
- plot_dendrogram.R - dendrograms

vis_ - specific plotting functions organized by topic 
- vis_hvf.R - highly variable features plots
- vis_pc.R - principle components plots
- vis_spatial.R 
- vis_spatial_in_situ.R

gg_ - ggplot2 related
- gg_info_layers.R - modular internals for plotting the different layers of information


Other:
- color_palettes.R - color palettes and factory functions


Package:
- package_imports.R - all imports
- globals.R - global variables
- dd.R - dummy documentation for inheriting
- zzz.R - onloads


## Spatial plots

Load the package

```{r}
library(GiottoVisuals)
```

Create a small Giotto object

```{r}
expression_matrix <- matrix(rnorm(1000), nrow = 10)
colnames(expression_matrix) <- paste0("cell_", 1:100)
rownames(expression_matrix) <- paste0("gene_", 1:10)

spatial_locations <- data.frame(
    cell_ID = paste0("cell_", 1:100),
    sdimx = rnorm(100),
    sdimy = rnorm(100)
)

giotto_object <- GiottoClass::createGiottoObject(
    expression = expression_matrix,
    spatial_locs = spatial_locations
)
```

Plot the spatial locations

```{r}
spatPlot2D(giotto_object)
```


## Session Info

```{r}
sessionInfo()
```

